---
title: "Oxford Two-Mode Policy Lifecycle REM Diagnostic"
author: "Jingyi Kang"
date: "June 2026"
output:
  html_document:
    toc: true
    toc_depth: 3
  pdf_document:
    toc: true
    toc_depth: 3
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

required_packages <- c(
  "data.table", "dplyr", "tidyr", "readr", "haven",
  "lubridate", "knitr"
)

for (pkg in required_packages) {
  suppressPackageStartupMessages(library(pkg, character.only = TRUE))
}

find_project_root <- function() {
  input_file <- tryCatch(knitr::current_input(), error = function(e) NULL)
  input_dir <- if (!is.null(input_file) && nzchar(input_file) && file.exists(input_file)) {
    dirname(normalizePath(input_file, mustWork = TRUE))
  } else {
    normalizePath(getwd(), mustWork = TRUE)
  }

  candidates <- unique(normalizePath(c(
    input_dir,
    getwd(),
    file.path(getwd(), "phd_work_rem_project"),
    dirname(getwd())
  ), mustWork = FALSE))

  for (candidate in candidates) {
    if (
      file.exists(file.path(candidate, "phd_work_rem_project.Rproj")) &&
        dir.exists(file.path(candidate, "data_raw", "oxford"))
    ) {
      return(normalizePath(candidate, mustWork = TRUE))
    }
  }

  stop("Could not locate phd_work_rem_project.Rproj and data_raw/oxford.")
}

project_root <- find_project_root()
knitr::opts_knit$set(root.dir = project_root)
path <- function(...) file.path(project_root, ...)

dir.create(path("derived"), showWarnings = FALSE, recursive = TRUE)
dir.create(path("outputs"), showWarnings = FALSE, recursive = TRUE)
dir.create(path("figures"), showWarnings = FALSE, recursive = TRUE)

fmt_pct <- function(x, digits = 1) paste0(round(100 * x, digits), "%")
```

# Project Check

```{r project-check}
project_check <- tibble::tibble(
  item = c(
    "Project root",
    "RStudio project file exists",
    "Oxford raw-data folder exists",
    "Derived-output folder",
    "Model-output folder"
  ),
  value = c(
    project_root,
    file.exists(path("phd_work_rem_project.Rproj")),
    dir.exists(path("data_raw", "oxford")),
    path("derived"),
    path("outputs")
  )
)

project_check |>
  knitr::kable()
```

# Purpose

This document focuses on one model track only: a REM-like event diagnostic for
policy lifecycle transitions in a two-mode country-policy process.

The two modes are:

- countries;
- policy indicators.

The event is a state transition in a country-policy pair:

$$
e_m = (country_i, policy_p, transition_r, time_t).
$$

The purpose is not to claim a final causal diffusion design. The purpose is to
test a narrower methodological question:

> Do transition-specific risk sets and eligibility-aligned peer exposure
> denominators change the interpretation of policy diffusion event models?

# Scope For This Sprint

This file estimates one part only: the event-based lifecycle diagnostic. A
snapshot-based network model would require a separate aggregation decision, so
it is not mixed into this script.

The immediate aim is to make the event definitions, risk sets, exposure
denominators, and missing-data choices explicit enough to discuss and revise.

# Raw Data

The Oxford/OxCGRT data were copied into this project under `data_raw/`. These
raw files should be treated as read-only. All derived data and outputs are
written to `derived/` and `outputs/`.

```{r raw-data-status}
oxford_files <- file.path(
  path("data_raw", "oxford"),
  paste0("OxCGRT_fullwithnotes_national_", 2020:2022, "_v1.csv")
)

cepii_file <- path("data_raw", "covariates", "dist_cepii.dta")

raw_status <- tibble::tibble(
  file = c(oxford_files, cepii_file),
  exists = file.exists(file),
  writable = file.access(file, 2) == 0,
  size_mb = round(file.info(file)$size / 1024^2, 1)
)

raw_status |>
  mutate(file = basename(file)) |>
  knitr::kable()
```

# Missing-Data And Coding Decisions

The missing-data strategy is part of the model definition because it affects
which country-policy-days can enter a transition risk set.

```{r missing-data-decisions}
missing_strategy <- tibble::tribble(
  ~decision_point, ~current_default, ~reason, ~sensitivity_needed,
  "Current policy value is missing",
  "Exclude from transition-event coding for that day.",
  "Missing is not the same as policy absence.",
  "Check whether missingness is concentrated in particular countries, indicators, or periods.",
  "Previous policy value is missing",
  "Treat as left-censored stock; do not code as ordinary adoption.",
  "The first observed positive value may reflect a policy already in force before the observation window.",
  "Run a sensitivity check where first observed positive values are counted as adoption.",
  "Current value is zero after a positive previous value",
  "Code as termination.",
  "The policy moves from active to inactive.",
  "Check whether zero sometimes means unreported rather than inactive.",
  "Current value decreases but remains positive",
  "Code as relaxation.",
  "The policy remains active but becomes less intense.",
  "Check whether ordinal steps are comparable across indicators.",
  "Peer exposure denominator",
  "Use both naive peer exposure and eligibility-aligned peer exposure.",
  "Peers that could not experience the same transition should not necessarily enter the denominator.",
  "Compare coefficient interpretation across the two exposure definitions."
)

missing_strategy |>
  knitr::kable()
```

Current default:

1. Do not recode missing policy values as zero.
2. Require both current and previous policy values to be observed for transition
   modelling.
3. Treat first observed positive values as left-censored initial stock, not as
   ordinary adoption.
4. Report a later sensitivity check that counts first observed positive values
   as adoption.

# Policy Indicators

```{r policy-indicators}
indicator_domains <- c(
  "C1E_School closing" = "Containment",
  "C2E_Workplace closing" = "Containment",
  "C3E_Cancel public events" = "Containment",
  "C4E_Restrictions on gatherings" = "Containment",
  "C5E_Close public transport" = "Containment",
  "C6E_Stay at home requirements" = "Containment",
  "C7E_Restrictions on internal movement" = "Containment",
  "C8E_International travel controls" = "Containment",
  "E1_Income support" = "Economic",
  "E2_Debt/contract relief" = "Economic",
  "H1_Public information campaigns" = "Health",
  "H2_Testing policy" = "Health",
  "H3_Contact tracing" = "Health",
  "H6E_Facial Coverings" = "Health",
  "H7_Vaccination policy" = "Health",
  "H8E_Protection of elderly people" = "Health",
  "V1_Vaccine Prioritisation (summary)" = "Vaccine",
  "V2A_Vaccine Availability (summary)" = "Vaccine",
  "V3_Vaccine Financial Support (summary)" = "Vaccine",
  "V4_Mandatory Vaccination (summary)" = "Vaccine"
)

tibble::tibble(
  policy = names(indicator_domains),
  domain = unname(indicator_domains)
) |>
  knitr::kable()
```

# Build Two-Mode Country-Policy-Day Panel

```{r load-and-build-panel}
read_oxford_year <- function(file) {
  available <- names(data.table::fread(file, nrows = 0))
  wanted <- c(
    "CountryName", "CountryCode", "Jurisdiction", "Date",
    "ConfirmedCases", "ConfirmedDeaths",
    names(indicator_domains)
  )
  keep <- intersect(wanted, available)
  dt <- data.table::fread(file, select = keep, showProgress = FALSE)
  for (missing_col in setdiff(wanted, names(dt))) {
    dt[, (missing_col) := NA]
  }
  dt[, source_file := basename(file)]
  dt[]
}

ox_raw <- data.table::rbindlist(lapply(oxford_files, read_oxford_year), fill = TRUE)
ox_raw <- ox_raw[Jurisdiction == "NAT_TOTAL"]
ox_raw[, Date := lubridate::ymd(as.character(Date))]

long <- data.table::melt(
  ox_raw,
  id.vars = c("CountryName", "CountryCode", "Date", "ConfirmedCases", "ConfirmedDeaths"),
  measure.vars = names(indicator_domains),
  variable.name = "policy",
  value.name = "policy_value"
)

long[, policy_value := suppressWarnings(as.numeric(policy_value))]
long[, domain := unname(indicator_domains[as.character(policy)])]
data.table::setorder(long, CountryCode, policy, Date)

long[, current_observed := !is.na(policy_value)]
long[, previous_value := data.table::shift(policy_value), by = .(CountryCode, policy)]
long[, previous_observed := !is.na(previous_value)]
long[, previous_row_exists := seq_len(.N) > 1L, by = .(CountryCode, policy)]
long[, previous_missing_due_first_row := !previous_row_exists]
long[, previous_missing_due_prior_missing := previous_row_exists & !previous_observed]
long[, policy_max_observed := max(policy_value, na.rm = TRUE), by = policy]
long[!is.finite(policy_max_observed), policy_max_observed := NA_real_]
long[, active := as.integer(current_observed & policy_value > 0)]
long[, previous_active := as.integer(previous_observed & previous_value > 0)]

long[, event_type := data.table::fcase(
  current_observed & previous_observed & previous_value == 0 & policy_value > 0,
  "adoption",
  current_observed & previous_observed & previous_value > 0 & policy_value > previous_value,
  "strengthening",
  current_observed & previous_observed & previous_value > 0 & policy_value < previous_value & policy_value > 0,
  "relaxation",
  current_observed & previous_observed & previous_value > 0 & policy_value == 0,
  "termination",
  current_observed & previous_observed,
  "persistence",
  current_observed & !previous_observed & policy_value > 0,
  "left_censored_initial_stock",
  current_observed & !previous_observed & policy_value == 0,
  "initial_absence",
  default = "missing_current"
)]

long[, reduction := as.integer(event_type %in% c("relaxation", "termination"))]
long[, risk_adoption := current_observed & previous_observed & previous_value == 0]
long[, risk_strengthening := current_observed & previous_observed & previous_value > 0 & previous_value < policy_max_observed]
long[, risk_reduction := current_observed & previous_observed & previous_value > 0]
long[, model_observed_lag := current_observed & previous_observed]
long[, month := format(Date, "%Y-%m")]

readr::write_csv(as.data.frame(long), path("derived", "country_policy_day_panel.csv"))

panel_summary <- tibble::tibble(
  countries = dplyr::n_distinct(long$CountryCode),
  policies = dplyr::n_distinct(long$policy),
  country_policy_days = nrow(long),
  observed_lagged_rows = sum(long$model_observed_lag),
  left_censored_initial_stock = sum(long$event_type == "left_censored_initial_stock"),
  missing_current_rows = sum(long$event_type == "missing_current")
)

panel_summary |>
  knitr::kable()
```

# Missing-Data Frequency Check

This check separates three issues that should not be mixed together:

1. current policy values that are missing;
2. lagged values that are missing only because the row is the first observation
   for that country-policy pair;
3. lagged values that are missing because the previous recorded value was
   missing.

```{r missing-frequency-check}
missing_overall <- tibble::tibble(
  measure = c(
    "Country-policy-days",
    "Current policy value missing",
    "Previous value missing: first row only",
    "Previous value missing: prior recorded value missing",
    "First observed positive value",
    "Rows usable for lagged transition modelling"
  ),
  count = c(
    nrow(long),
    sum(!long$current_observed),
    sum(long$previous_missing_due_first_row),
    sum(long$previous_missing_due_prior_missing),
    sum(long$event_type == "left_censored_initial_stock"),
    sum(long$model_observed_lag)
  ),
  percent_of_country_policy_days = c(
    1,
    mean(!long$current_observed),
    mean(long$previous_missing_due_first_row),
    mean(long$previous_missing_due_prior_missing),
    mean(long$event_type == "left_censored_initial_stock"),
    mean(long$model_observed_lag)
  )
)

missing_by_domain <- long[, .(
  rows = .N,
  current_missing = sum(!current_observed),
  current_missing_pct = mean(!current_observed),
  prior_value_missing = sum(previous_missing_due_prior_missing),
  prior_value_missing_pct = mean(previous_missing_due_prior_missing),
  first_observed_positive = sum(event_type == "left_censored_initial_stock"),
  first_observed_positive_pct = mean(event_type == "left_censored_initial_stock"),
  usable_lagged_rows = sum(model_observed_lag),
  usable_lagged_rows_pct = mean(model_observed_lag)
), by = domain][order(domain)]

missing_by_policy <- long[, .(
  rows = .N,
  current_missing = sum(!current_observed),
  current_missing_pct = mean(!current_observed),
  prior_value_missing = sum(previous_missing_due_prior_missing),
  prior_value_missing_pct = mean(previous_missing_due_prior_missing),
  first_observed_positive = sum(event_type == "left_censored_initial_stock"),
  first_observed_positive_pct = mean(event_type == "left_censored_initial_stock"),
  usable_lagged_rows = sum(model_observed_lag),
  usable_lagged_rows_pct = mean(model_observed_lag)
), by = .(policy, domain)][order(-current_missing_pct, policy)]

readr::write_csv(missing_overall, path("outputs", "missing_overall.csv"))
readr::write_csv(as.data.frame(missing_by_domain), path("outputs", "missing_by_domain.csv"))
readr::write_csv(as.data.frame(missing_by_policy), path("outputs", "missing_by_policy.csv"))

missing_overall |>
  mutate(percent_of_country_policy_days = fmt_pct(percent_of_country_policy_days, 3)) |>
  knitr::kable()

missing_by_domain |>
  mutate(
    current_missing_pct = fmt_pct(current_missing_pct, 3),
    prior_value_missing_pct = fmt_pct(prior_value_missing_pct, 3),
    first_observed_positive_pct = fmt_pct(first_observed_positive_pct, 3),
    usable_lagged_rows_pct = fmt_pct(usable_lagged_rows_pct, 3)
  ) |>
  knitr::kable()

missing_by_policy |>
  mutate(
    current_missing_pct = fmt_pct(current_missing_pct, 3),
    prior_value_missing_pct = fmt_pct(prior_value_missing_pct, 3),
    first_observed_positive_pct = fmt_pct(first_observed_positive_pct, 3),
    usable_lagged_rows_pct = fmt_pct(usable_lagged_rows_pct, 3)
  ) |>
  knitr::kable()
```

For the current proof of concept, missing values are treated as missing rather
than inactive policy states. The key empirical check is whether missingness is
rare and whether it is concentrated in particular policy domains or indicators.

# Event And Risk-Set Summary

```{r event-risk-summary}
event_counts <- long[
  event_type %in% c("adoption", "strengthening", "relaxation", "termination", "left_censored_initial_stock"),
  .(events = .N),
  by = event_type
][order(event_type)]

risk_counts <- data.table::data.table(
  transition = c("adoption", "strengthening", "reduction"),
  risk_rule = c(
    "previous value = 0",
    "previous value > 0 and below policy-specific maximum",
    "previous value > 0"
  ),
  risk_rows = c(
    sum(long$risk_adoption, na.rm = TRUE),
    sum(long$risk_strengthening, na.rm = TRUE),
    sum(long$risk_reduction, na.rm = TRUE)
  ),
  events = c(
    sum(long$event_type == "adoption", na.rm = TRUE),
    sum(long$event_type == "strengthening", na.rm = TRUE),
    sum(long$reduction == 1, na.rm = TRUE)
  )
)
risk_counts[, event_rate := events / risk_rows]

readr::write_csv(as.data.frame(event_counts), path("outputs", "event_counts.csv"))
readr::write_csv(as.data.frame(risk_counts), path("outputs", "risk_counts.csv"))

event_counts |>
  knitr::kable()

risk_counts |>
  mutate(event_rate = fmt_pct(event_rate, 3)) |>
  knitr::kable()
```

The important point is not the absolute count alone. The denominator changes
when the event is adoption, strengthening, or reduction. A pooled model that
treats all observed country-policy-days as equally at risk would mix different
logical opportunities.

# Peer Network

The current peer network uses CEPII geography/language information:

- contiguity;
- common official language;
- inverse log distance.

This is a placeholder network for the proof of concept. It should later be
compared with other pathways such as trade, ideology, regional organisation
membership, or prior policy similarity.

```{r peer-network}
countries <- sort(unique(long$CountryCode))

cepii <- as.data.table(haven::read_dta(cepii_file))
cepii <- cepii[iso_o %in% countries & iso_d %in% countries & iso_o != iso_d]
cepii[, inv_log_dist := 1 / log1p(as.numeric(dist))]
cepii[, raw_weight := 1.5 * as.numeric(contig) + 0.75 * as.numeric(comlang_off) + inv_log_dist]
cepii[is.na(raw_weight), raw_weight := 0]

W <- matrix(0, nrow = length(countries), ncol = length(countries), dimnames = list(countries, countries))
for (i in seq_len(nrow(cepii))) {
  W[cepii$iso_o[i], cepii$iso_d[i]] <- cepii$raw_weight[i]
}
diag(W) <- 0
row_sums <- rowSums(W)
W[row_sums > 0, ] <- W[row_sums > 0, ] / row_sums[row_sums > 0]

peer_network_summary <- tibble::tibble(
  countries = length(countries),
  countries_with_positive_peer_weight = sum(rowSums(W) > 0),
  directed_positive_edges = sum(W > 0),
  density = sum(W > 0) / (length(countries) * (length(countries) - 1))
)

readr::write_csv(as.data.frame(peer_network_summary), path("outputs", "peer_network_summary.csv"))
peer_network_summary |>
  knitr::kable()
```

# Exposure Construction

For each country-policy-day, the focal exposure is recent peer reduction on the
same policy indicator.

Two exposure definitions are compared:

1. **Naive peer reduction exposure**: weighted count of peer reduction events in
   the previous 14 days.
2. **Eligibility-aligned peer reduction rate**: weighted peer reduction events
   divided by weighted peer-days that were eligible to reduce in the previous
   14 days.

```{r exposure-construction}
rolling_lag_sum <- function(mat, window = 14) {
  out <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  if (nrow(mat) <= 1) return(out)
  for (lag in seq_len(window)) {
    if ((lag + 1) <= nrow(mat)) {
      out[(lag + 1):nrow(mat), ] <- out[(lag + 1):nrow(mat), ] + mat[1:(nrow(mat) - lag), ]
    }
  }
  out
}

all_dates <- sort(unique(long$Date))
rem_parts <- list()

for (p in names(indicator_domains)) {
  sub <- long[as.character(policy) == p]
  model_rows <- sub[model_observed_lag == TRUE]
  if (nrow(model_rows) == 0) next

  event_mat <- matrix(0, nrow = length(all_dates), ncol = length(countries), dimnames = list(as.character(all_dates), countries))
  risk_mat <- matrix(0, nrow = length(all_dates), ncol = length(countries), dimnames = list(as.character(all_dates), countries))

  event_rows <- sub[reduction == 1 & !is.na(Date) & CountryCode %in% countries]
  if (nrow(event_rows) > 0) {
    event_mat[cbind(as.character(event_rows$Date), event_rows$CountryCode)] <- 1
  }

  risk_rows <- sub[risk_reduction == TRUE & !is.na(Date) & CountryCode %in% countries]
  if (nrow(risk_rows) > 0) {
    risk_mat[cbind(as.character(risk_rows$Date), risk_rows$CountryCode)] <- 1
  }

  lag_event <- rolling_lag_sum(event_mat, 14)
  lag_risk <- rolling_lag_sum(risk_mat, 14)

  peer_event <- lag_event %*% t(W)
  peer_risk <- lag_risk %*% t(W)

  model_rows[, date_pos := match(as.character(Date), rownames(peer_event))]
  model_rows[, country_pos := match(CountryCode, colnames(peer_event))]
  idx <- cbind(model_rows$date_pos, model_rows$country_pos)

  model_rows[, naive_peer_reduction_14 := as.numeric(peer_event[idx])]
  model_rows[, eligible_peer_mass_14 := as.numeric(peer_risk[idx])]
  model_rows[, eligible_peer_reduction_rate_14 := fifelse(
    eligible_peer_mass_14 > 0,
    naive_peer_reduction_14 / eligible_peer_mass_14,
    0
  )]
  model_rows[, date_pos := NULL]
  model_rows[, country_pos := NULL]

  rem_parts[[p]] <- model_rows
}

rem_data <- data.table::rbindlist(rem_parts, fill = TRUE)
rem_data[, reduction := as.integer(reduction)]

readr::write_csv(as.data.frame(rem_data), path("derived", "rem_exposure_panel.csv"))

exposure_summary <- rem_data[, .(
  rows = .N,
  reduction_events = sum(reduction),
  mean_naive_peer_reduction_14 = mean(naive_peer_reduction_14, na.rm = TRUE),
  mean_eligible_peer_rate_14 = mean(eligible_peer_reduction_rate_14, na.rm = TRUE),
  mean_eligible_peer_mass_14 = mean(eligible_peer_mass_14, na.rm = TRUE)
)]

readr::write_csv(as.data.frame(exposure_summary), path("outputs", "exposure_summary.csv"))
exposure_summary |>
  knitr::kable(digits = 4)
```

# Model Comparison

Three specifications are compared.

| Model | Focal rows | Exposure |
|---|---|---|
| A | all observed country-policy-days after lag construction | naive peer reduction count |
| B | only reduction-eligible country-policy-days | naive peer reduction count |
| C | only reduction-eligible country-policy-days | eligibility-aligned peer reduction rate |

The models are deliberately simple. They are diagnostic models, not final causal
diffusion models.

```{r model-comparison}
sample_for_logit <- function(dt, event_col = "reduction", ratio = 20, seed = 20260614) {
  set.seed(seed)
  events <- dt[get(event_col) == 1]
  nonevents <- dt[get(event_col) == 0]
  n_nonevents <- min(nrow(nonevents), max(nrow(events) * ratio, 1))
  nonevents_sample <- nonevents[sample(.N, n_nonevents)]
  data.table::rbindlist(list(events, nonevents_sample), fill = TRUE)
}

fit_one_logit <- function(data, formula, model_name) {
  fit <- glm(formula, data = as.data.frame(data), family = binomial())
  co <- summary(fit)$coefficients
  coefficients <- data.frame(
    model = model_name,
    term = rownames(co),
    estimate = co[, 1],
    std_error = co[, 2],
    p_value = co[, 4],
    odds_ratio = exp(co[, 1]),
    row.names = NULL
  )
  summary_row <- data.frame(
    model = model_name,
    rows = nrow(data),
    events = sum(data$reduction),
    event_rate = mean(data$reduction),
    aic = AIC(fit),
    bic = BIC(fit)
  )
  list(coefficients = coefficients, summary = summary_row)
}

model_a_data <- rem_data[model_observed_lag == TRUE]
model_b_data <- rem_data[risk_reduction == TRUE]
model_c_data <- model_b_data

model_a_sample <- sample_for_logit(model_a_data, seed = 1)
model_b_sample <- sample_for_logit(model_b_data, seed = 2)
model_c_sample <- sample_for_logit(model_c_data, seed = 3)

model_a <- fit_one_logit(
  model_a_sample,
  reduction ~ previous_value + naive_peer_reduction_14 + domain + month,
  "A_naive_denominator"
)

model_b <- fit_one_logit(
  model_b_sample,
  reduction ~ previous_value + naive_peer_reduction_14 + domain + month,
  "B_state_dependent_denominator"
)

model_c <- fit_one_logit(
  model_c_sample,
  reduction ~ previous_value + eligible_peer_reduction_rate_14 + eligible_peer_mass_14 + domain + month,
  "C_state_denominator_eligible_exposure"
)

model_summary <- dplyr::bind_rows(model_a$summary, model_b$summary, model_c$summary)
model_coefficients <- dplyr::bind_rows(model_a$coefficients, model_b$coefficients, model_c$coefficients)

readr::write_csv(model_summary, path("outputs", "rem_model_summary.csv"))
readr::write_csv(model_coefficients, path("outputs", "rem_model_coefficients.csv"))

model_summary |>
  mutate(event_rate = fmt_pct(event_rate, 2), aic = round(aic, 2), bic = round(bic, 2)) |>
  knitr::kable()

model_coefficients |>
  filter(term %in% c(
    "naive_peer_reduction_14",
    "eligible_peer_reduction_rate_14",
    "eligible_peer_mass_14"
  )) |>
  mutate(
    estimate = round(estimate, 4),
    std_error = round(std_error, 4),
    p_value = signif(p_value, 3),
    odds_ratio = round(odds_ratio, 3)
  ) |>
  knitr::kable()
```

# Interpretation

This comparison should be read as a diagnostic. It asks whether the model
changes when the risk set and peer-exposure denominator are aligned with the
focal transition.

The result should not be described as proof that rollback diffuses. The current
design still needs:

1. domestic severity controls;
2. population denominators for cases and deaths;
3. alternative peer networks;
4. uncertainty clustered by country and policy;
5. a non-Covid benchmark.

# Variables Or Decisions Still Needed

```{r variables-needed}
needed <- tibble::tribble(
  ~item, ~why_it_matters, ~current_status,
  "Population denominator",
  "Confirmed cases/deaths should be converted to per-capita severity measures.",
  "Not included in this first REM diagnostic.",
  "Lagged severity controls",
  "Needed to reduce common-shock confounding in Covid policy changes.",
  "Oxford contains ConfirmedCases and ConfirmedDeaths; the lagged per-capita construction still needs to be agreed.",
  "Alternative peer network",
  "CEPII geography/language is only one possible diffusion pathway.",
  "Need to add trade, regional organisations, ideology, policy similarity, or another defensible pathway.",
  "Missing-value sensitivity",
  "First observed positive values may affect adoption and stock counts.",
  "Default is left-censoring; sensitivity check still needed.",
  "Non-Covid benchmark",
  "Required because Covid may be unusual.",
  "IEA or a policy diffusion replication package should be checked next."
)

needed |>
  knitr::kable()
```

# Six-Month Output Logic

```{r six-month-plan}
six_month_plan <- tibble::tribble(
  ~period, ~deliverable,
  "Month 1", "Definitive event-coding and missing-data memo; clean two-mode Oxford panel; sensitivity check for first observed positive values.",
  "Month 2", "REM-like lifecycle diagnostic comparing naive denominator, state-dependent denominator, and eligibility-aligned exposure.",
  "Month 3", "Improved covariates and peer network: severity controls, population denominator, and at least one non-geographic exposure network.",
  "Month 4", "Robustness checks: alternative exposure windows, alternative missing-data coding, and clustered uncertainty.",
  "Month 5", "Non-Covid benchmark check using IEA or a policy diffusion replication package.",
  "Month 6", "Focused methods memo with reproducible code appendix and decision on the main methodological contribution."
)

six_month_plan |>
  knitr::kable()
```

# Conclusion

This REM-only version is the cleanest immediate proof of concept because it
keeps the event process in daily time and directly addresses the main
methodological issue from the report: policy lifecycle transitions require
transition-specific risk sets.

The strongest claim at this stage is:

> Risk-set construction and peer-exposure denominator choices are inferential
> choices in policy lifecycle diffusion models, not merely coding choices.

The claim remains diagnostic and preliminary. It does not yet solve causal
identification.
